CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-09-25 PaperRMFI2019/samplepaper.tex
add EN standard

The SMPL model is at its top level a skinned articulated 
model, i.e., 
consists of a 
surface mesh $\mathcal{M}$ that mimics the skin and a corresponding skeleton 
$\mathbf{J}$. The mesh $\mathcal{M} \in \mathbb{R}^{3N}$, which is a boundary 
representation stores 
both the body geometry (vertex position) and topology (how the vertices are 
connected). The skeleton $\mathbf{J}$ is defined by its endpoints (joints) 
location in 3D space $j_i \in \mathbb{R}^3$  and its kinematic tree. Two 
connected 
joints define a 'bone' $P_i$. Moreover, a child bone is rotated relative to its 
connected parent. The pose $\vec{\theta}$ is the specification of every bone 
rotation in angle-axis form plus an orientation for the root bone.

More specifically, the SMPL model is defined by a template mesh (mean template 
shape) $\mathbf{\bar{T}}$ represented by a vector of $N = 6890$ concatenated 
vertices in the zero pose, $\vec{\theta}^*$. The skeleton has $K = 23$ joints, 
therefore a pose vector is defined by $|\vec{\theta}| = 3 \times 23 + 3 = 72$ 
parameters. In order 
to synthesize a new human mesh one has to deform the 
provided template mesh by 
setting shape parameters $\vec{\beta}$. Pose parameters $\vec{\theta}$ are used 
for animation. 
The 
model provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}.
\end{equation}

@@

Recall that mesh $\mathcal{M}$ has $N$ vertices. Let the set of vertices be 
$\mathcal{V}$ and $j^y_i$ the y-coordinate of vertex $j_i$, we can segment the 
mesh in chest, waist and pelvic regions 
$\mathcal{CR}, \mathcal{WR}, \mathcal{PR}$, respectively

\begin{align}
\mathcal{CR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

\begin{align}
\mathcal{WR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

\begin{align}
\mathcal{PR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla based on 
the mesh signature. We could then establish the location index as lower bound 
for the search of chest location. There is consensus regarding the axilla 
definition as the area of the human body underneath the shoulder joint 
(\cite{MeSH.axilla}, \cite{FMA.axilla}, \cite{TA.axilla}). Therefore, we can 
use the shoulder joint as a hint to locate the axilla and establish a lower 
bound for the chest region. We then can combine this idea with the defintion 
used in the garment industry, namely the maximum horizontal girth under the 
armpits (axillae) (rh).
Let us consider the left shoulder joint on in our skeleton model $J_{ls}$. We 
cast a ray in direction $d$ outside the mesh. Since the joint $J_{ls}$ lies 
inside the mesh, the ray intersects its surface at point $p_{axilla}$ (see 
figure x).
With this construction we can define the chest region $\mathcal{CR}$ as the 
region where the slice index is bigger that the index of a slice with 
y-coordinate is equal or greater than the axilla $y$-coordinate 
$y_{p_{axilla}}$.

\begin{align}
\mathcal{CR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

Similarly, the waist region is

\begin{align}
\mathcal{WR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

Finally, the pelvic region is

\begin{align}
\mathcal{PR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. This intuition is compliant 
with prior body dimensions standardized definition, for example the European 
standard EN 13402-1. Similarly, the pelvis 
circumference is measured often around the widest part of hips and 
buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at a point $j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the intersection boundary length as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, j) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $j_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters until we reach the bounding box bottom 
$j_b$. Because the mesh $\mathcal{M}$ and the plane normal $\vec{n}_{||}$ 
remains the same we can drop them from the notation. For every slice at point 
$j$, we compute $\mathcal{BL}(j)$.

\begin{align}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ \mathcal{BL}(j_t), 
\mathcal{BL}(j_t+m), \cdots, \mathcal{BL}(j_b) \right]
\end{align}

Finally, we can define the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ that maps every slice index $k \in \{0, 1, \cdots, 
|\mathcal{L}|-1\}$ to the corresponding boundary length $\mathcal{BL}(j)$ in 
the mesh slide vector $\vec{\mathcal{L}}$.

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(k) = \mathcal{BL}(j)
\end{align}

@@

\begin{figure}[h]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

>>>>

The SMPL model is at its top level a \textbf{skinned articulated 
model}, i.e., 
consists of a 
surface mesh $\mathcal{M}$ that mimics the skin and a corresponding skeleton 
$\mathbf{J}$. The mesh $\mathcal{M} \in \mathbb{R}^{3N}$, which is a boundary 
representation stores 
both the body geometry (vertex position) and topology (how the vertices are 
connected). The skeleton $\mathbf{J}$ is defined by its joints location in 3D 
space $j_i \in \mathbb{R}^3$  and its kinematic tree. Two 
connected 
joints define a `bone'. Moreover, a child bone is rotated relative to 
its 
connected parent. The pose $\vec{\theta}$ is the specification of every bone 
rotation plus an orientation for the root bone.

The SMPL model is also a \textbf{deformable model} 
\cite{Terzopoulos:1987:EDM:37401.37427}. In order to 
synthesize a 
new human mesh one has to deform the provided template mesh by 
setting shape parameters $\vec{\beta}$. Pose parameters $\vec{\theta}$ are used 
for animation.

More specifically, the model is defined by a template mesh (mean template 
shape) $\mathbf{\bar{T}}$ represented by a vector of $N = 6890$ concatenated 
vertices in the zero pose, $\vec{\theta}^*$. The skeleton has $K = 23$ joints.

As an \textbf{statistical model}, SMPL provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}.
\end{equation}

@@

Let the axilla location be the point $p = (p^x, p^y, p^z)$ and 
$j_{nw}$, $j_{pl}$, 
$j_{lh} \in \mathbf{J}$ be the natural waist, pelvis and left hip joints, 
respectively.

Recall that mesh $\mathcal{M}$ has $N$ vertices. Let the set of vertices be 
$\mathcal{V}$ and $v^y_i$ the y-coordinate of vertex $v_i$, we can 
segment the 
mesh in chest, waist and pelvic regions 
$\mathcal{CR}, \mathcal{WR}, \mathcal{PR}$, respectively

\begin{align}\label{eq:cr}
\mathcal{CR} = \big(v^y_i \, | \, p^y > v^y_i \geq j^y_{nw} \big)
\end{align}

\begin{align}\label{eq:wr}
\mathcal{WR} = \big(v^y_i \, | \, j^y_{nw} > v^y_i \geq j_{pl}\big)
\end{align}

\begin{align}\label{eq:pr}
\mathcal{PR} = \big(v^y_i \, | \, j_{pl} > v^y_i \geq j_{lh}\big)
\end{align}

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point in the mesh at the proper location under 
the arms. One point suffices because we require only a reference along the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the left shoulder joint $j_{ls} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray in direction $d$ outside the mesh. The 
joint $j_{ls}$ lies inside the mesh thus the ray intersects its surface at 
point $p$ (see 
figure x).

Once we have a segmented mesh, we search the dimensions within 
these regions. The body dimensions we want to calculate are 
geodesic on the mesh. Next subsection discuss how.

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference bellow the arms 
at the \textit{widest} part of the torso and the waist circumference at the 
\textit{narrowest} part after the chest but above the hips. This intuition is 
compliant 
with prior body dimensions standardized definition, for example, the European 
standard EN 13402-1 \cite{en13402-1}. Similarly, the pelvis 
circumference is measured often around the widest part of hips and 
buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at a point $q \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the intersection boundary length as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters until we reach the bounding box bottom 
$q_b$. Because the mesh $\mathcal{M}$ and the plane normal $\vec{n}_{||}$ 
remains the same we can drop them from the notation. For every slice at point 
$q$, we compute $\mathcal{BL}(q)$.

\begin{align}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ \mathcal{BL}(q_t), 
\mathcal{BL}(q_t+m), \cdots, \mathcal{BL}(q_b) \right]
\end{align}

Finally, we can define the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ that maps every slice index $k \in \{0, 1, \cdots, 
|\mathcal{L}|-1\}$ to the corresponding boundary length $\mathcal{BL}(q)$ in 
the mesh slide vector $\vec{\mathcal{L}}$.

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(k) = \mathcal{BL}(q)
\end{align}

@@

\begin{figure}[h!]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}
